---
title: "SEC 2-FA2 GROUP 24-VALDEZ, CJ"
subtitle: "https://github.com/ChanceValdez/APM1110/tree/main/FA2"
author: "Chance Jayden Valdez"
date: "August 30, 2026"
output: pdf_document
---

## Item 4
Amy and Jane are gambling together. A fair coin is tossed repeatedly. Each time a head comes up, Amy wins two euro from Jane, and each time a tail comes up, Amy loses two euro to Jane. Use R to simulate this game 100 times:

**R Code to Simulate Tossing Game**
```{r}
num_tosses <- 100
H <- 2
T <- -2

tosses <- sample(c(H, T), size = num_tosses, replace = TRUE)
winnings <- cumsum(tosses)
plot(1:num_tosses, winnings, type = "o", 
     xlab = "Number of Tosses", ylab = "Winnings")
abline(h = 0)
```
Now, estimate:

(a) the number of times that Amy is ahead in these 100 tosses;

(b) how much Amy has won or lost after 100 tosses.

**R Code to Answer (a) and (b)**

```{r}
times_ahead <- sum(winnings > 0)
cat("The number of times Amy was ahead in these", num_tosses, "tosses:\n")
cat("Amy was ahead:", times_ahead, "times.\n")

final_winnings <- winnings[num_tosses]
cat("How much Amy has won or lost after", num_tosses, "tosses:\n")
cat("Amy's final net winnings/losses:", final_winnings, "euros.\n")
```

**Additional Calculations (for fun!)**

```{r}
times_tied <- sum(winnings == 0)
times_behind <- sum(winnings < 0)

cat("Amy was tied (at 0 euros):", times_tied, "times.\n")
cat("Amy was behind:", times_behind, "times.\n")
```

**Github:** https://github.com/ChanceValdez/APM1110/tree/main/FA2
